### 2019-06-25 main/draft.tex
Before redoing the introduction

\documentclass[12pt,letter]{article}
\usepackage[left=0.8in,right=0.8in,top=1in,bottom=1in]{geometry}
\usepackage{amsmath}
\usepackage{pdflscape}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{microtype}
\usepackage{euscript}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{mathrsfs}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\tikzstyle{block} = [rectangle, draw, rounded corners]
\tikzstyle{line} = [draw, -latex']
\newcommand{\hypo}{\mathcal{H}}            
\bibliographystyle{ieeetr}

\usepackage[flushleft]{threeparttable}

%\usepackage[cp1251]{inputenc}
\usepackage[english]{babel} 
\DeclareMathOperator{\rank}{rank}
\newcommand*{\hm}[1]{#1\nobreak\discretionary{}
            {\hbox{$\mathsurround=0pt #1$}}{}}

            \def\onepc{$^{\ast\ast}$} \def\fivepc{$^{\ast}$}
\def\tenpc{$^{\dag}$}
\def\legend{\multicolumn{4}{l}{\footnotesize{Significance levels
:\hspace{1em} $\dag$ : 10\% \hspace{1em}
$\ast$ : 5\% \hspace{1em} $\ast\ast$ : 1\% \normalsize}}}

\newcommand{\bs}[1]{\boldsymbol{#1}}  
\newcommand{\bsA}{\boldsymbol{A}}

%\setstretch{1}                         
\flushbottom                            
\righthyphenmin=2                      
\pagestyle{plain}                       
%\settimeformat{hhmmsstime}  
\widowpenalty=300                   
\clubpenalty=3000                     
\setlength{\parindent}{0em}           
\setlength{\topsep}{0pt}              
\usepackage[pdftex,unicode,colorlinks=true,urlcolor=blue]{hyperref}
\usepackage{bbm}
\usepackage{tabularx}
\renewcommand{\emptyset}{\varnothing}

\setlength{\parskip}{0.5\baselineskip plus2pt minus2pt}

\newcommand{\e}{\varepsilon}
\DeclareMathOperator*{\Argmax}{\mathrm{Argmax}}
\DeclareMathOperator*{\Argmin}{\mathrm{Argmin}}
\DeclareMathOperator*{\argmax}{\mathrm{arg\,max}}
\DeclareMathOperator*{\argmin}{\mathrm{argmin}}

\newcommand{\blp}{\mathrm{BLP}}
\DeclareMathOperator*{\plim}{\mathrm{plim}}
\DeclareMathOperator{\Max}{\mathrm{Max}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\Z}{\mathcal{Z}}
\renewcommand{\geqslant}{\geq}
\renewcommand{\leqslant}{\leq}
\newcommand{\p}{\bs p}
\newcommand{\y}{\bs y}
\def\dd#1#2{\frac{\partial#1}{\partial#2}}

\renewcommand{\emptyset}{\varnothing}

\DeclareMathOperator{\tr}{\mathrm{tr}}

\newcommand{\bb}{\bs \beta}
\newcommand{\X}{\bs X}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\PP}{\mathbb{P}}
\DeclareMathOperator{\V}{\mathbb{V}}
\DeclareMathOperator{\CM}{\mathbb{C}}
\renewcommand{\C}{\CM}
\DeclareMathOperator{\var}{\mathrm{var}}
\DeclareMathOperator{\cov}{\mathrm{cov}}
\DeclareMathOperator{\corr}{\mathrm{corr}}
\DeclareMathOperator{\MSE}{\mathrm{MSE}}
\DeclareMathOperator{\Bias}{\mathrm{Bias}}
\renewcommand{\P}{\PP}
\newcommand{\dsim}{\stackrel{d}{\sim}}
\newcommand{\hn}{\mathcal{H}_0}
\newcommand{\ha}{\mathcal{H}_a}
\newcommand{\thetab}{\bs \theta}
\newcommand{\pv}{\text{P-value}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\MLE}{\scriptscriptstyle MLE}
\newcommand{\LR}{\mathrm{LR}}
\newcommand{\I}{\mathbb{I}}
\newcommand{\sumin}{\sum\limits_{i=1}^n}
\newcommand{\sumti}{\sum\limits_{t=0}^\infty}
\newcommand{\hbeta}{\hat{\beta}}
\newcommand{\halpha}{\hat{\alpha}}
\newcommand{\hsigma}{\hat{\sigma}}
\newcommand{\hvar}{\widehat{\var}}
\newcommand{\hcov}{\widehat{\cov}}
\newcommand{\Q}{\mathbb{Q}}

\newcommand{\pconv}{\xrightarrow{ \ p \ }}
\newcommand{\dconv}{\xrightarrow{ \ d \ }}
\newcommand{\asconv}{\xrightarrow{ \ a.s. \ }}
\newcommand{\msconv}{\xrightarrow{ \ m.s. \ }}

\newcommand{\pic}[4][h!]{\begin{figure}[#1]

\begin{center}\includegraphics[width=#2cm]{#3}\caption{#4\label{#3}}\end{center}
\end{figure}}

%outtex
\def\onepc{$^{\ast\ast}$} \def\fivepc{$^{\ast}$}
\def\tenpc{$^{\dag}$}
\def\legend{\multicolumn{4}{l}{\footnotesize{Significance levels
:\hspace{1em} $\dag$ : 10\% \hspace{1em}
$\ast$ : 5\% \hspace{1em} $\ast\ast$ : 1\% \normalsize}}}
%end outtex

%\bibliographystyle{ieeetr}

\newcommand{\laseq}{\stackrel{\lambda\text{-a.e.}}{=}}
\renewcommand{\d}{\underline}
\renewcommand{\u}{\overline}
\newcommand{\td}{\underline{\theta}}
\newcommand{\tu}{\overline{\theta}}
%\renewcommand{\theenumi}{\alph{enumi}}
%\renewcommand{\labelenumi}{(\theenumi)}
%\renewcommand{\theenumii}{\roman{enumii}}
%\renewcommand{\labelenumii}{\theenumii.}
%\renewcommand{\theenumiii}{\arabic{enumiii}}
%\renewcommand{\labelenumiii}{\theenumiii.}
%\renewcommand{\epsilon}{\varepsilon}
\newcommand{\hneq}{\stackrel{\hn}{=}}
\newcommand{\deq}{\stackrel{d}{=}}
\title{Kids First or Marriage First?}
\author{Egor Kozlov}
\begin{document}
%\begin{center}\textbf{Economics 416-1} \\ \emph{By Egor Kozlov}\end{center}
\maketitle

\begin{abstract}
Unplanned pregnancies often trigger marriage for unmarried couples. I document two robust empirical facts. First, women who have their first child before or at the year of marriage are statistically more likely to be divorced in the future than those who marry before they have a child. Second, this difference is more pronounced for women with higher income and educational attainment, although it is less common for them to have their kids before marriage. I estimate a lifecycle model with dynamic bargaining, marriage selection and unplanned pregnancy risks. I show that the first fact can be attributed to household bargaining friction, as women agree to enter lower quality marriages to avoid being a single mother, and  the second fact is consistent with childcare requiring both time and money, as this suggests that the costs of being a single mother are higher for more productive women. Qualitatively, this sheds some light on household formation and dissolution and importance of accounting for heterogeneous time costs. Qualitatively, the model suggests sizable causal effect of unplanned pregnancies on observable divorce rates.
\end{abstract}

\section{Introduction}
As of 2018, [40\%] of new births occurred by unmarried mothers. This number is rising for decades, but, more importantly, it is rather heterogeneous, reaching three-quarters for mothers who are African American, [...] under 25 and [...] for high school or less in education. But even for educated and older women this share is substantial [...]. An increase in non-marital cohabitation is also a documented trend, despite of this, persistent non-marital cohabitation is still not commong: according to National Survey of Family And Growth, the median duration of cohabitation is 22 months, and median duration of cohabitation if it ends by marriage is 21.\footnote{Source: Copen et al, \cite{copen}}. Unmarried unions with children are still pretty rare: according to Payne, 2013\nocite{payne}, only about 3\% of children live in unions with unmarried cohabitating parents (as opposed to 21\% with single mothers).

These numbers have important economic implications: children born by unmarried women, as documented by, for instance Kearney, Levine, 2017, display significantly worse future outcomes in terms of education and income, and this pattern survives controlling for many dimensions. Although it is well understood that children of never married single mothers or children of blended families (when a woman marries someone else than child's father) have generally worse outcomes, even when parents marry each other after they made a baby their marriage is statistically less stable, as I argue in this paper, and the larger chances of marriage dissolution do affect future prospects of the children.

This pairs well with common perception that some couples marry at or shortly after the moment they make a baby. If cohabitation is viewed as uncommitted relationship, unexpected fertility may serve as a trigger to obtain more commitment. Narrowly, we can interpret this as shotgun marriage story, however in the times with large abortion access many unplanned pregnancies are aborted, and cohabitating couples seem to \emph{choose} to keep an unplanned baby. The whole role of marriage changes as it is more socially acceptable to cohabitate, though it is much more conventional to do it before having kids. As highlighted by, for instance, Lafortune, 2018, despite of general decrease of role of marriage in the society, securing investment to children is one of the most important forces that makes people marry each other.

This paper highlights one important aspect of this: I argue that when marriage decision is driven by having an unexpected child, bargaining friction makes marital outcome worse. The source of this friction is the large childcare costs and unequal relative exposure to them: as women bear relatively more childcare costs, they are more likely to enter relatively lower quality marriages as a commitment to share a part of these costs. This can rationalize observable differences in marital outcomes between couples with and without out-of-wedlock births.

I document that females who had their first birth before their first marriage have higher chances of divorce in the future, this is mainly driven by women married to the parents of their children, and this is robust and relatively more pronounced for high education or high income groups. I then provide a model incorporating the bargaining mechanism that can rationalize these differences. I estimate the model parameters to match observable patterns in the data and show what it implies about the role of unequal exposure to childcare costs in observable marriage selection and divorce patterns. Shortly, the model suggests if people could perfectly control timing of their births, general share of divorced people would decrease by [...], especially in high-income groups.

Section 2 of this paper discuss existing results and general trends that are already described by the literature. Section 3 discuss how I construct the dataset and timing variables, and Section 5 establishes main observations about differences in groups depending on their marriage timing, together with a number of additional considerations useful to understand the content of these differences. After that I explain theoretical model in Section 5 and the process of setting its parameters in Section 6. Section 7 describes estimation results and counterfactual simulations implied by these results. Section 8 provides more general discussion and concludes.

\section{Related Literature}

My work builds on few branches of literature, following long tradition of modeling fertility and marriage markets in family economics. I am not aware of any studies that cover precisely the narrow issue I focus on, but there are few recently established result that cover related aspects and compliment what I do. Three main building blocks that I utilize are literature on marriage selection and dynamic bargaining, few detailed lifecycle models of fertility and unplanned pregnancies and number of descriptive studies of microdata describing reduced form effects of out-of-wedlock births, trends in divorce rates and heterogeneity in time use of parents.

The main conceptual framework that I utilize are models of household's dynamic bargaining and limited commitment, extensively reviewed by Chiaporri, Mazzocco 2015.\nocite{chiappori-review} These models were used to answer various marriage and divorce related questions. Theoretical part of my model is mostly based on Voena, 2015\nocite{voena-1}, that uses dynamic bargaining to consider impact of changes in divorce laws, and Low et al, 2018\nocite{low-1}, focusing on marriage and divorce selection related to welfare utilitzation. There are few more recent papers utilitizing similar approaches, for instance, Shephard, 2019\nocite{shephard} studying gender and age pay gaps through equilibrium on marriage markets.

To model fertility choices and shocks, I rely on recent lifecycle models. I make many choices similar to Sommer, 2015\nocite{sommer} who studies how changes in reproductive horizon affect couple's fertility decisions with in the environment with income shocks, paying detailed attention to time and money structure of childcare. More detailed treatment of unintended pregnancies is done by Ejrnæs, Jørgensen, 2018\nocite{ejrnaes}, they extend the model to allow for non-random abortion decisions though do not model childcare costs in details.

Commitment in marriage and its interaction with fertility decisions attract recent attention as possible drivers of trends in marriage and divorce. Work by Lafortune, Low 2018\nocite{lafortune-1} uses ACS data and housing prices, as well as marriage market model, to argue that married people buy houses as a device to commit not to divorce, that allows them to secure investment in their future children. This is parallel and complimentary with my results, that mainly focus on consequences of not having enough commitment and how they relate to fertility timing.

Impact of non-marital childbearing and parents divorce are well-documented empirically from children prospective. For example, Kearney, Levine 2017 use PSID to document how being born out of marriage statistically child's outcome in school years and in early adult life, and the heterogeneity of this by parent's education and age groups. Shortly, they find presence of ``marriage premium'' for children at each education level, though for less educated and younger mothers benefits of having married parents are more pronounced in early life, and for elder and more educated mothers --- in adult life. A classic study by Akerlof, 1996\nocite{akerlof} highlight the mechanism why unplanned pregnancies may not be aborted in couples so they are still present even with large abortion access. Finally, Ginther, Pollak, 2004 show that outcomes of children in blended families (i.e. having step-parents) are similar to outcomes of children of single parents.

This paper fills an important gap in the literature considering impact of out-of-marriage births for women. As fertility, marriage and divorce involve many individual decisions affected by many unobservable factors, it is not possible to ask casual questions by using only the data, and existing empirical studies only provide descriptive evidence. This study contributes to the growing literature of quantitative models that both use the tools of family economics to endogenize timing of fertility, marriage, divorce and their interactions, and, on the other hand, provide numbers and patterns consistent with microdata, so they can be used to answer policy questions.

As noted by Ginther and Pollak, causal questions in the context of marriage and divorce should be taken with caution: to answer, for instance, how parents' divorce affect child's future we have to be precise what counterfactual we compare it with (either parents never divorced, parents stayed together for few more years, parents never married). The same issues concern impact of unplanned pregnancies in cohabitating couples, therefore, instead of trying to ask causal quesitons I the main goal of using the model is to deliver counterfactual outcomes, that still facilitate understanding of some causality.

\section{Empirical Facts}

This section presents key empirical facts and several exercises on testing their validity. In summary:
\begin{enumerate}
\item Women who had their first child before their first marriage (K$\to$M) are more likely to be divorced than those who had their first child in their first marriage (M$\to$K).
\item Relative difference in the share of divorced between K$\to$M and M$\to$K is increasing with income and education
\item These results are statistically significant and robust to various methods of controlling for observable composition of the groups
\end{enumerate}

To establish these results, I mainly use American Community Survey (ACS), that has large sample sizes and many relevant observable characteristics of households. Its only limitation is that it does not recover marriage and fertility history perfectly. Because of this, I re-confirm the results with the data from Survey of Income and Program Participation (SIPP 2014), that has fewer observations but more detailed marriage and fertility history.

I begin with describing empirical strategy, then Section \ref{keycomp} presents the key results for ACS and SIPP, the following sections present several validity checks.

\subsection{Defining Groups}

General strategy involves dividing sample in two groups: couples who had kids before their marriage ($K\to M$) and those who had kids after their marriage ($M\to K$). After this division I perform multiple comparisons of these groups: the main comparison is looking at divorce rates in several variations, few additional comparisons involve marriage inequality and outcomes of children.

Although the couples are the primary focus of the paper, the surveys I use look at individuals and households. Therefore I follow females rather than males: upon divorce, children are more likely to stay with mothers (share of single mothers in ACS is ..., single fathers ...). The timing I consider in the  most general form is (for females, given both events occurred): 
\[\Delta T = T^{\text{First child  is born}} - T^{\text{First marriage happened}},\]
where $T$ refers to moments of time. After that I define two groups (given $\Delta T$ is measured in years):
\begin{align*}&M\to K \Leftrightarrow \Delta T \in \{1,2,...\},\\
&K\to M \Leftrightarrow \Delta T \in \{...,-2,-1,0\},\end{align*}

In practice, I exclude observations with $\Delta T < -5$ or $\Delta T > 10$. The reason for the first one is mainly related to step-children: for the subsample when it can be tracked, share of step-children is becomes higher with more negative $\Delta  T$ in ACS, and having step-children that are relatively adult at the moment of marriage is a complex issue that is beyond the scope of the paper. See a note ... in the appendix about this. The reason for the second one is mainly data quality (...).

As the evidence is mainly cross-sectional, I refer to share of divorced people in a group as divorce rates. Unlike fertility, people can change their marriage status all the time, so there is no such thing as completed divorce rate, therefore I just use the share of people who ever had a divorce. More accurate measuring of divorce rates involves conditioning on age, however fixing age restricts the attention to a particular cohort, and pooling people of different ages together allows to think about more general population.

\subsection{Sample Restrictions}
\subsubsection{ACS data}
I take ACS samples of 2008--2016, as only they have variables concerning the duration of current marital status. Namely, people report the year of their most recent marriage. Fertility history is not observed perfectly: the survey reports age of the eldest child residing with individual. However, both of these variables can recover the timing of interest for relatively large subpopulation of females who:
\begin{enumerate}
\item Had exactly one marriage, so they are either married or divorced now (treating legally separated, spouse absent and widowed as married)
\item Have children present in household, with age of the eldest child below 18
\item Most likely reside with their eldest child judging by mother's age (I pick 21--40)
\end{enumerate}
%The last restriction can be relaxed, but it allows to track partners and children more clearly and also insures better quality of self-reported data.

The most important restriction here is being married once --- that means that people who re-marry drop out of sample. For instance, if it take longer to re-marry for the people who had kids before their first marriage, share of divorced people in the group $K\to M$ would be larger without differences in probability of their first marriage to dissolve. This concern can only partially be addressed with ACS data, but this is not an issue in SIPP.

All mean and quantile calculations are performed using person weights.

\subsubsection{SIPP data}

I use wave 1 of 2014 SIPP. It it representative for the whole population, I use person weights in all calculations.

SIPP has much more detailed information on marital history and fertility: it explicitly asks about times of first marriage and first births, it also asks about multi-partner fertility. However it has substantially smaller sample size. For SIPP I relax the restrictions and consider all females who had at least one marriage and are of age 20--40 (treating married with spouses absent, separated and widowed as married), as their marital and fertility history is recovered perfectly.

%When dividing people by income, however, I have to further restrict attention to households that consist of mother + spouse/partner (if reported) + children. This creates some attrition especially for young people, but differences in defining household members are minimal between ACS and SIPP, so it is not likely that it creates an issue.

\subsection{Income Partition\label{inc-part}}
I define income groups based on per hour income of working females, by excluding those who does not work or work too little from income considerations. Precisely, I define working as those who are employed now, report to work more than 25 weeks in the last year, report non-missing labor income above or equal to $\$1000$ per year, report non-missing usual working hours above or equal to $5$ hours per week.

As weeks worked last year are defined only on intervals for ACS, I impute them using the midpoints of respective intervals, then define yearly hours as a product of usual hours and imputed weeks worked. This allows to define log per-hour income as log-difference of yearly log wage and salary income with yearly log hours. I residualize these income data by regressing it on 4th degree polynomial in age, dummies and linear trends in age for each US state and dummies for each ACS survey year.

In SIPP 2014, I perform the similar operations. I exclude females who had less than $1000/12$ monthly income or had average weekly hours less than 5 during the last month. Then I divide monthly income on four times average weekly hours to get estimate for per-hour earnings. As before, I residualize logs of these income by regressing them on the same factors, except that instead of year controls (I only use 2014) I use month controls.

\subsection{Raw Comparisons\label{keycomp}}
The main results I discuss is summarized by Figure \ref{shrs} for ACS and Figure \ref{shrs-sipp} for SIPP. Table \ref{diff-raw} in presents these results in numbers.

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\linewidth]{shrs.eps}
\begin{tabular}{|p{0.65\linewidth}|}
\hline
\footnotesize $\Delta T = \text{Year}^{\text{first child}} - \text{Year}^{\text{first marriage}}$. Restriction: $\Delta T \in \{-5,...,10\}$\\
\footnotesize Definition: K$\to$M if $\Delta T \in \{-5,...,0\}$. M$\to$K if $\Delta T \in \{1,...,10\}$.\\\hline
\end{tabular}
\caption{Share of K$\to$M couples and share of divorced, ACS\label{shrs}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\linewidth]{shrs-sipp.eps}
\caption{Share of K$\to$M couples and share of ever divorced, SIPP\label{shrs-sipp}}
\end{figure}

\begin{table}[p]
\begin{center}
\begin{tabular}{|l|c||c|c||c||c|c|}\cline{2-7}
\multicolumn{1}{c|}{} & \multicolumn{3}{c||}{\small ACS: sample of married once} & \multicolumn{3}{c|}{\small SIPP: sample of  ever married} \\\cline{2-7}
\multicolumn{1}{c|}{}  & \multirow{2}{*}{\small \% K$\to$M}& \multicolumn{2}{c||}{\textbf{\% divorced (now)}} & \multirow{2}{*}{\footnotesize \% K$\to$M }& \multicolumn{2}{c|}{\textbf{\% divorced (ever)}} \\\cline{3-4}\cline{6-7}
\multicolumn{1}{c|}{}  &  &\small if M$\to$K &  \small if K$\to$M & & \small if M$\to$K & \small if K$\to$M \\\hline
\textit{All females, 21--40}  & $23.2\%$ & $10.1\%$ & $18.0\%$ & $22.8\%$ & $24.1\%$ & $34.8\%$ \\
\textit{All females, 35--40} & $18.0\%$ & $11.3\%$ & $23.0\%$ & $18.9\%$ & $28.1\%$ & $45.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Education partition.}}\\\hline
\textit{High school only, 21--40} & $34.2\%$ & $13.9\%$ & $17.2\%$ & $33.4\%$ & $29.2\%$  & $34.2\%$ \\
\textit{College graduates, 21--40} & $9.9\%$ & $5.3\%$ & $14.7\%$ & $10.5\%$ & $12.9\%$ & $27.5\%$ \\\hline\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize \textbf{Income partition:} per-hour income for subsample of working women.}\\\hline
\textit{Low 30\% income, 21--40} & $30.8\%$ & $17.6\%$ & $24.1\%$ & $34.6\%$ & $30.5\%$ &  $42.5\%$ \\
\textit{Mid 40\% income, 21--40} & $25.5\%$ & $14.2\%$ & $21.5\%$ & $24.1\%$ & $28.3\%$ & $45.8\%$ \\
\textit{Top 30\% income, 21--40} & $14.1\%$ & $7.7\%$ & $18.1\%$ & $12.4\%$ & $22.9\%$  & $32.8\%$ \\\hline
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize $\Delta T = \text{Year}^{\text{first child}} - \text{Year}^{\text{first marriage}}$. Restriction: $\Delta T \in \{-5,...,10\}$.}\\
\multicolumn{7}{|p{0.9\linewidth}|}{\footnotesize Definition: K$\to$M if $\Delta T \in \{-5,...,0\}$. M$\to$K if $\Delta T \in \{1,...,10\}$.}\\
\hline
\end{tabular}
\caption{Differences in raw means, ACS and SIPP (repeats Figures \ref{shrs} and \ref{shrs-sipp}) \label{diff-raw}}
\end{center}
\end{table}

Three key observations from the table are:
\begin{enumerate}
\item About one-fifth of females have their first kid before or at the year they marry, this share quickly decreases with income and education
\item Share of divorced people among those who had their first kid before on at the year of their first marriage is larger.
\item The share of divorced people is about 1.5 higher for low income/education females and more than twice higher for high income/education.
\item The ACS result is not likely to be driven by picking only people with one marriage, as the pattern is exactly the same if I use SIPP and count for remarriages
\end{enumerate}

This alone is not enough to draw strong conclusions: composition of these groups is drastically different. For instance, as suggested by the table, people with lower income are more likely to get divorced, so even conditional on income/education group K$\to$M couples may be poorer with each subgroup. Additionally, many other factors may drive the differences, including different ages of children (people in K$\to$M group by construction have older babies after any given number of years after marriage), regional and racial differences in attitudes to divorce, age composition, possible presence of step-children in K$\to$M couples and so on. I attempt to control for few of these things and argue that they explain these differences only partially, that suggests the presence of causal mechanism that I describe.

\subsection{Adjusting for Composition}
As people in K$\to$M and M$\to$K groups obviously have different characteristics, as in particular suggested by Figure \ref{keycomp}, I use regression techniques to see what part of difference is attributed to this composition. Appendix \ref{comp-diff-appendix} quantifies these differences, shortly, people in K$\to$M group are younger, less educated, have kids earlier, marry slightly earlier and have more unequal marriages. In this part I control for difference in observable characteristics to see what part of the difference is attributed to the fact of being in K$\to$M group. Shortly, controlling for observables mitigate some part of the difference but it still remains sizable and significant.

In general, I perform the following exercise: I regress indicator of female being divorced on set of controls $X$ and dummy variable $\I(\text{K$\to$M})_i$. The specification I use is
\[D_i = \Delta \cdot \I(\text{K$\to$M})_i + X_i\gamma + \varepsilon_i.\]
Replacing $X_i$ with a constant corresponds to raw mean comparison described above.

I report three possible estimates of $\Delta$:
\begin{itemize}
\item \textbf{Raw}, corresponding to differences in raw means in the previous part
\item \textbf{Controls}, where the regression includes dummies for ACS year, the US state, number of children, number of years after marriage, age of the eldest child, race and education group + 4th degree polynomial in woman's age
\item \textbf{Controls+income}, where the regression includes all of the above + 3rd degree polynomial in log female per-hour income (residualized as described above) for subsample of working women. This subsample is smaller, but has arguably better data quality.
\end{itemize}

See Table \ref{diff-contr} for the results. Generally, controlling for composition decreases observable difference, but this decrease is less than half in magnitude: even controlling for many observable factors, many of which may itself be results of unplanned fertility, we cannot assign all observable difference to composition of the groups. This also implicitly suggests that the income partition I choose makes sense: conditional on income group, additional controlling for income does not change the observable coefficient much.

\begin{table}
\begin{center}
\begin{tabular}{|l|c||c|c|c|c|c|c|}\cline{2-8}
\multicolumn{1}{c|}{} & \multicolumn{7}{c|}{ACS: sample of married once}\\\cline{2-8}
\multicolumn{1}{c|}{} & \small Share of &\multicolumn{2}{c|}{\small Raw difference} & \multicolumn{2}{c|}{\small with controls} & \multicolumn{2}{c|}{\small controls + income} \\ \cline{3-8}
\multicolumn{1}{c|}{} & \small divorced & $\Delta$ & (s.e.) & $\Delta$ & (s.e.) & $\Delta$ & (s.e.) \\  \hline
\textit{All females, 20--50} & 14.3\% &  7.9\% &   (0.1\%) &    5.7\%  &    (0.2\%) &     4.6\%   &   0.2\% \\
\textit{All females, 35--40} &  17.3\% &  11.7\%  &  (0.2\%) &   6.7\% &  (0.3\%) &     6.2\%  &   (0.4\%) \\\hline\hline
\multicolumn{8}{|p{0.9\linewidth}|}{\footnotesize \textbf{Education partition.}}\\\hline
\textit{High school only, 20--50} & 18.1\% &   3.3\%  &    (0.2\%) &   8.4\% &   (0.3\%)   &  7.8\%   &  (0.5\%) \\
\textit{College graduates, 20--50} & 8.5\% &   9.4\% &   (0.2\%) &   3.9\% &   (0.3\%) &   3.2\%   &  (0.4\%) \\\hline\hline
\multicolumn{8}{|p{0.9\linewidth}|}{\footnotesize \textbf{Income partition:} working women only.}\\\hline
\textit{Low 30\% income, 20--40} &  21.6\% &  6.5\% &    (0.3\%) &    7.8\% &  (0.5\%) &   7.8\%   &  (0.5\%) \\
\textit{Mid 40\% income, 20--40} &  17.1\% &  7.3\% &   (0.3\%) &    3.6\%   & (0.4\%) &   3.6\% &    (0.4\%) \\
\textit{Top 30\% income, 20--40} & 11.1\% &  10.4\% &  (0.3\%) &    3.5\% &   (0.4\%) & 3.5\% &  (0.4\%) \\
\hline
\end{tabular}
\caption{Differences, controlled for composition, ACS\label{diff-contr}}
\end{center}
\end{table}

\subsection{Additional Comparisons}

Appendix \ref{extra-comparisons} shows few additional graphs and exercise. I summarize their results here:
\begin{enumerate}
\item In the sample I use M$\to$K women are married for longer at the moment of survey\footnote{This does not contradict the fact that K$\to$M couples marry slightly earlier as I use the sample of married women.} (so they had \emph{more} years to break up), comparing marriages of the same duration only exacerbates the difference.
\item Children of K$\to$M women are on average older (as they are born in earlier ages). Divergence between K$\to$M and M$\to$K observable divorce probabilities happens around ages 5--10. After controlling for composition, however, I show that K$\to$M couples divorce more often even conditional on child's age.
\item To further address concerns about duration, I recompute the numbers using getting the divorce at the year of survey, corresponding to marginal divorce probability as opposed to cumulative divorce probability (being divorced at by the moment survey). The magnitudes and patterns are very similar, though the numbers are harder to interpret. This suggests that 
\item I show that the difference is not likely to be driven by presence of stepchildren in K$\to$M families. In fact, as I restrict $\Delta T \geq -5$, share of stepchildren is relatively small, and surviving couples with stepchildren do not differ in observable marriage duration with couples without stepchildren.
\end{enumerate}

\subsection{Intra-Household Allocations [TBD]}
I document few additional facts describing the difference between K$\to$M and M$\to$K besides being divorced. Following Lafortune et al, 2018, I construct measure of children's grade retention based on ACS (that is an indicator if any school-age child in the household is in grade that is lower than prescribed by her/his age). I show this together with few other intra-houeshold characteristics: measure of inequality and labor force participation.

See Table \ref{intra} for the numbers. The exercise suggests that children on K$\to$M women repeat grades significantly more often, and this happens even in higher-income households. [This table is based on old sample so it is slightly inconsistent with what I do above, I will redo this in one of the future iterations.]

\begin{table}
\centering
\begin{tabular}{|l||c|c||c|c|}\hline
& \multicolumn{2}{|c||}{\footnotesize \textit{All couples}}  & \multicolumn{2}{|c|}{\footnotesize \textit{Top 40\% of income} } \\\hline
&                       \footnotesize  M$\to$K & \footnotesize  K$\to$M &  \footnotesize  M$\to$K & \footnotesize  K$\to$M \\\hline
\footnotesize \textit{Female in Labor Force}        & $70.3\%$  & $71.4\%$  & $79.3\%$ & $85.3\%$  \\\hline
\footnotesize \textit{Earninigs Ratio}    &  $95.2\%$ & $93.9\%$  &  $95.8\%$ &  $96.4\%$ \\\hline
\footnotesize \textit{Kids' Grade Retention}  & $4.6\%$    & $7.6\%$ &   $3.3\%$   & $5.9\%$ \\\hline
\end{tabular}
\caption{Child's outcomes and intra-household allocations [to be fixed later]\label{intra}}
\end{table}

%\section{Mechanisms}
%This part describes the mechanism I am trying to capture in the model that is presented in the following section. Suppose single agents with characteristics $\omega^m$ and $\omega^f$ can meet each other and form a couple. On top of individual characteristics, let couples be characterized by idiosyncratic utility gain $\psi$ from being together, that can be interpreted as a love shock. Let also there exists a variable $\theta$ that describes terms of marriage (that can be interpreted as decision weight of wife in a collective household). Let for $i \in \{f,m\}$ functions $V^{i,c}$ and $V^{i,s}$ denote values of being in a couple and being single, respectively. People agree to marry if there exist marriage terms $\theta$ such that gains of both spouses from marriage is positive, that is
%\[\Delta^f = V^{f,c}(\omega^f,\omega^m,\theta,\psi) - V^{f,s}(\omega^f) \geq 0, \ \ \Delta^m = V^{m,c}(\omega^f,\omega^m,\theta,\psi) -  V^{m,s}(\omega^m) \geq 0.\]
%
%Assume that utilities of potential spouses increase in $\psi$. Assume also that wife benefits from higher $\theta$ and husband benefits from lower $\theta$. The following illustration captures the intuition for the model:

\section{Model}

\input{model}

\section{Quantification}
I use techniques of indirect inference to fit the model to the data. Namely, given the parameter values I solve for decision rules and simulate $5{,}000$ agents for the first 20 periods of their life corresponding to ages of (21--40 used in the data part). During estimation procedure I look for the parameter values such that observable differences between between the simulated and actual data, measured by the targets I describe below, is minimal.

\subsection{Externally Set Parameters}
Length of the period is set to be 1 year, the model assumes people start their life with no assets at the age of $21$ and die deterministically at the age of $81$ (so $T = 60$). However, it assumes that fertility status can change only till age of 37 ($T_f = 17$) and marriage status till age of 45 ($T_m = 25$). It also assumes that after the age of 60 people retire and get $b = 0.4$ as in Sommer, 2014 of their income at age of 60 (so no transitions happen after $T = 40$). I only simulate the data for age 21--40, therefore these choices are not likely to play a crucial role.

Following standard approach in the macro literature, I choose $\sigma = 1.5$, $R = 1.03$ and $\beta = 1/R$. Also, following estimates of Voena, 2014 I set returns to scale in couple's consumption to be $\rho_c = 0.23$.

I take age-income profiles  $(a^f_0,a^f_1,a^f_2)$, $(a^m_0,a^m_1,a^m_2)$ to match coefficients in per-hour labor earnings in cross-sectional ACS regressions. I borrow variances of income shocks $\sigma^{z,f}$, $\sigma^{z,m}$ from Low et al 2018 as they cannot be estimated correctly in cross-sectional data, and although they were estimated for subsample of non college graduates in SIPP, in the model they produce inter-quaritle ratio at the age of 30 that is very similar to the full sample of ACS data.

I also externally estimate $\sigma^{z,p}$ as variance of (unpredicted part of) log-difference in spouses earnings, see Appendix \ref{partearn} for this. Based on this calculations I set $\sigma^{z,p} = 0.6$ (this reproduces $\sigma = 0.7$ in the simulated data because income shocks are uncorrelated in the model so they are likely to diverge over time). As I both do not have good assets data and do not aim to replicate the assets distribution, I set $\sigma^{a,p} = \sigma^{z,p}$.

I also set $p_s = \frac14$ to roughly reflect that children go to school on average at 6 and to pre-school few years earlier and this should correspond to a change in childcare costs structure.

In the part with assets, I assume savings can be between 0 and 20 times average labor income in the beginning of life. This bar in pretty low, but it is enough to create reasonable assets variation at the moment most of marriage decisions are made, creating median assets-to-income ratio at 30 being slightly above 1.

$9\%$ of individuals are assumed to be married at the age of $21$ ($t = 1$) as suggested by ACS, no one is assumed to have any children at $t = 1$ for simplicity. People start their life with zero assets and with productivity distribution as in Low et al 2018 (variances of initial $z$ are $0.15$ for women and $0.18$ for men).

Random walk variables are limited to stay within $\pm 20$ one-period standard deviations around zero. Marriage terms $\theta$ are forced to be within $[0.1,0.9]$ as wider bounds occasionally create numerical issues.

\subsection{Estimated Parameters}
This section describes parameters estimated jointly in simulated method of moments, as well as targets that are relevant for capturing each parameter. Shares of divorced for K$\to$M and M$\to$K groups are used in addition to these targets, it is not obvious which parameters they identify directly, but they help to overcome possibly insufficient identifying power of the ones listed below.

\begin{table}[h!!]
\begin{tabular}{|c|p{0.25\linewidth}|c|p{0.25\linewidth}|}\hline
Parameters & Description & Values & Identifying variation \\\hline
($\sigma_{\psi,0},\sigma_{\psi}$) & { \footnotesize
 Marriage surplus initial distribution \& shocks, standard deviations} &$(0.76,1.73)$ & {\footnotesize
 \% of divorced at $(25,35,40)$, \% of women with one marriage at 40} \\\hline\hline
$p^{\text{meet}}_{\{t = 21, 30, 37, 45\}}$ & {\footnotesize Meeting probabilities by age (linearly interpolated in between)} & 
 $(0.13,0.10,0.07,0.12)$ & {\footnotesize \% of never married by age , \% of women with one marriage at 40} \\\hline
$(\phi_{0,c},\phi_3)$ & {\footnotesize Scale parameters for utility of child's quality} & $(-1.89,2.58)$ & {\footnotesize \% of women with no children at 25, 35, and at 30 by income groups}\\\hline
$(\phi_1,\phi_2,\theta)$ & {\footnotesize Child quality preference structure and substitutability} & $(0.44,1.07,-1.02)$  & {\footnotesize Average share of money spent on children, ratio of money and time spent by income groups} \\\hline
$(\mu_y,\mu_a)$ & {\footnotesize Time structure of childcare by child's age} &  $(0.35,0.01)$ & {\footnotesize Time spent with children 0--5 relative to 6--12}  \\\hline
$\bar{q}$ & {\footnotesize Childcare quality requirement } & $2.12$ & {\footnotesize Income of never married with kids and divorced with kids relative to general population} \\\hline
$(p^{\text{out}},\phi_{0,s})$ & {\footnotesize Single mothers: probability of re-entering marriage market and utility shifter} & $(0.10,-3.34)$ & {\footnotesize \% of divorced with kids and never married with kids at $30$} \\\hline\hline
$(p^{\text{preg}}_0,p^{\text{preg}}_1,p^{\text{preg}}_2,p^{\text{preg}}_3)$  & {\footnotesize Arrival probabilities of unplanned pregnancies by age and $z$} & $(0.67,0.13,-0.50,-0.005)$ & {\footnotesize Share of K$\to$M at 25, 35; share of K$\to$M at 30 and 40 by income groups} \\\hline
\end{tabular}
\caption{Estimated parameters and relevant variation.}
\end{table}

%\subsection{Overview}
%Full list of parameters involves many things:
%\begin{itemize}
%\item Fundamental parameters of income fluctuation model $(\beta,\sigma,R)$
%\item Preferences for child's quality: $(\phi_0,\phi_1,\phi_2,\phi_3)$
%\item Parameters for child's quality production function: CES power $\theta_k$ and labor shares by age $(\mu_y,\mu_s)$
%\item Unavoidable childcare costs $\underline{q}$
%\item Transition probabilities:
%\begin{itemize}
%\item Probability of pregnancy shock by age: $p^{\text{preg}}_t = \lambda^{\text{preg}}_0 - \lambda^{\text{preg}}_1\cdot t$ (2 parameters)
%\item Probability of meeting a partner: $p^{\text{meet}}_t:$ piecewise linear function between $(p^{\text{meet}}_{21},p^{\text{meet}}_{27},p^{\text{meet}}_{34},p^{\text{meet}}_{40})$ (4 parameters)
%\item Probability of re-entering marriage market and corresponding utility ``compensation'' $(p^{\text{out}},S)$. I set $S = 0$ in current calibration (so 1 parameter).
%\item Probability that kids grow up (transition from $y$ to $s$): $p_s$ (1 parameter)
%\end{itemize}
%\item Age profiles of labor income by gender: $(a^f_0,a^f_1,a^f_2)$, $(a^m_0,a^m_1,a^m_2)$ (6 parameters)
%\item Variance of shocks (6 parameters):
%\begin{itemize}
%\item Individual income shocks $\sigma^{z,f}$, $\sigma^{z,m}$.
%\item Couple's surplus shocks $\sigma^{\psi}$
%\item Couple's initial surplus $\sigma^{\psi_0}$
%\item Assets and productivity of potential partner: $(\sigma^{z,p},\sigma^{a,p})$
%\end{itemize}
%\end{itemize}

%Few additional choices involve time horizons, bounds for random walk variables and bounds for assets and $\theta$.

%\subsection{Estimated Parameters}
%Out of 31, 15 parameters are set externally, so I am left with $16$ parameters to estimate: child's parameters $(\phi_0,\phi_1,\phi_2)$, child quality production parameters $(\theta_k,\mu_y,\mu_a)$, variances of initial marriage surplus and its innovations $(\sigma^{\psi_0},\sigma^{\psi})$, meeting probability parameters $(p^{\text{meet}}_{21},p^{\text{meet}}_{27},p^{\text{meet}}_{34},p^{\text{meet}}_{40})$, pregnancy probability parameters $(\lambda^{\text{preg}}_0,\lambda^{\text{preg}}_1)$, transition out of being single mother parameters $p^{\text{out}}$, $S$. Unfortunately, all of them interact and as the model is highly stylized but involves many choices, they cannot be estimated externally.

%Conceptually, they can be split onto following groups:
%\begin{enumerate}
%\item Unplanned pregnancy probabilities are pinned down by share of $K\to M$ females at ages 25 and 35, as well as share of planned births in total population
%\item Meeting probabilities and $\sigma^{\psi_0}$ are pinned down by shares of never married females people at 25, 35, 40 and share of females with one marriage at 40 (relevant to capture chances of re-marrying).
%\item  Preferences for kids are captured by share of childless females by age of 25, 30, 35 (identifies general scale $\phi_0$ and $\phi_3$), share of childless females for low and high income at 30 (captures $\phi_0$ and $\phi_1$). Luxury good parameter $\phi_2$ is mainly pinned down by ratio of amounts of money spent on kids between 75 and 25 income quantiles.
%\item Childcare production function is recovered by ratio of time spent with kids by couple's income (pins down elasticity of substitution), average share of money spent on kids (captures ``average'' $\mu$) and ratio of time spent by child's age (captures difference between $\mu_a$ and $\mu_y$)
%\item Variance of innovation to marriage surplus is mainly pinned down by share of divorced people by age (25, 35, 40).
%\item Transition out of single mothers is captured by share of never married single mothers at 25 and share of divorced women with kids at 30 (picks $p^{\text{out}}$).
%\end{enumerate}
%On top of it, I use shares of divorced people for $K\to M$ and $M \to K$ at 40 per se and by income groups as additional targets that are affected by mixture of parameters described above.
%

\subsection{Simulating Attrition\label{income-attr}}
For income considerations in the simulated data I put missing values of per-hour income for females/couples who have children and choose $l$ larger than its 58\% percentile. This corresponds to $58\%$ of ACS females of 21--40 who are employed and at least 5 hours and have children. I then perform similar detrending and divide individuals by group using the same regression procedure as described in \ref{inc-part}. This aims to replicate similar attrition in the data (as I focus on female's income many women with young children are missing from considerations), without explicitly modeling labor force participation margin.

In the same way, while classifying on M$\to$K and K$\to$M I exclude women who married more than once. I also exclude never married single mothers, even if they re-entered the marriage market and married someone. Therefore, in the model K$\to$M refers only to couples who got a pregnancy shock at the meeting ($\Delta T = 0$). A subtle consequence of this is that as I use K$\to$M share to discipline the pregnancy shock and assume that in existing couples probabilities of unplanned pregnancies are the same as for people who just met, inside-couple fertility may not be modeled correctly. In the current calibration, however, majority of couple's pregnancies are by choice rather that by fertility shock\footnote{This is consistent, with, for instance, \href{https://www.guttmacher.org/fact-sheet/unintended-pregnancy-united-states}{data from Guttmacher Institute} reporting that about 4 out of 10 births are unplanned.}, and this does not seem to affect the numbers significantly, so I currently avoid introducing another parameter that adjusts intensity of pregnancy shocks for married and unmarried women.

\section{Results}
Table \ref{targets-main} shows the targets in the model and in the data. The model fits reasonably well and replicated key magnitudes. [I keep updating the solution occasionally].

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|}\hline
Target & Model & Data \\\hline
\multicolumn{3}{|c|}{Marriage Dimension (source: my calculations on ACS)} \\\hline
Never married by 25  & $60.6\%$ & $68.3\%$ \\
Never married by 35 & $24.6\%$ & $25.6\%$  \\
Never married by 40 & $19.0\%$ & $18.8\%$  \\\hline
One marriage by 40  & $69.6\%$ & $64.7\%$ \\\hline
\multicolumn{3}{|c|}{Fertility Dimension (source: my calculations on ACS)} \\\hline
No kids at 25 & $65.0\%$  & $67.7\%$ \\
No kids at 30 (low $30\%$) & $52.1\%$  &   $44.0\%$ \\
No kids at 30 (top $30\%$) & $56.8\%$ &   $57.8\%$ \\
No kids at 35  &  $27.2\%$ &   $28.1\%$ \\
No kids at 40 & $23.8\%$ & $25.6\%$ \\\hline
\multicolumn{3}{|c|}{Marriage/Fertility Interactions (source: my calculations on ACS)} \\\hline
Never married with kids at 30 & $10.4\%$  & $14.1\%$ \\
Divorced with kids at 30 & $4.2\%$ & $4.0\%$ \\
{\footnotesize Income(Never mar with kids at 30) / Income(Everyone at 30)} & $86.5\%$ & $70.7\%$\\
{\footnotesize Income(Divorced with kids at 30) / Income(Everyone at 30)} & $82.2\%$ & $87.0\%$\\\hline
\multicolumn{3}{|c|}{Divorce Dimension } \\\hline
Divorced by 25 &  $4.5\%$  & $10.7\%$ \\
Divorced by 30 (low $30\%$)  & $17.2\%$ &     20.8\% \\
Divorced by 30 (top $30\%$)  & $9.0\%$ &     8.1\% \\
Divorced by 35  & $10.8\%$ &    15.9\% \\
Divorced by 40  & $16.3\%$  &    19.1\% \\\hline
\multicolumn{3}{|c|}{K$\to$M --- M$\to$K partition (source: my calculations on ACS)} \\\hline
Divorced by 40 if K$\to$M & 25.1\% &   23.5\% \\
Divorced by 40 if M$\to$K & 14.8\% &      13.0\% \\
Divorced by 40 if K$\to$M (low 30\%) & $32.5\%$ & 28.5\% \\
Divorced by 40 if K$\to$M (top 30\%) & $26.1\%$ & 23.3\% \\
Divorced by 40 if M$\to$K (low 30\%) & $23.0\%$ & 18.0\% \\
Divorced by 40 if M$\to$K (top 30\%) & $15.2\%$ & 11.3\% \\\hline
\multicolumn{3}{|c|}{Childcare (data: CEX and ATUS, sources in parenthesis)} \\\hline
Money for childcare /Total expenditures &$43.4\%$ & 40.0\% \cite{sommer} \\
Time spent 0--5 / Time spent 6--12 & $2.07$ & 2 [\href{https://www.bls.gov/charts/american-time-use/activity-by-parent.htm}{BLS}]\\
Time spent top 30\% / Time spent low 30\% & 1.11 & 1.19  \cite{schneider} \\
Money spent top 30\% / Money spent low 30\% & 2.56  & 2.67 \cite{schneider} \\\hline
\multicolumn{3}{|p{0.7\linewidth}|}{\footnotesize Note 1: shares of divorced are computed for subsample of married exactly once in both data and model.}\\
\multicolumn{3}{|p{0.7\linewidth}|}{\footnotesize Note 2: low 30\% -- mid 40\% -- top 30\% refers to woman's per hour income. The income partition is not full as income can be missing, see \ref{inc-part} and \ref{income-attr}.}\\\hline
\end{tabular}
\caption{Targets in the data and in the model\label{targets-main}}
\end{center}
\end{table}

\subsection{Counterfactual Simulations}

This part does two things. First, to understand the anatomy of result, I simulate individual impact of fertility shocks, comparing several counterfactual scenarios for simulated group of individuals. Second, I look at how changes in parameters driving fertility shocks and bargaining affect the observable (aggregate) moments.

\subsubsection{Individual Impact}

This part looks at counterfactual impact of the fertility shocks. To do this, I generate two sequences of shocks from the estimated distribution, $\mathcal{S}_{m,p}$ and $\mathcal{S}_{m,np}$. These sequences are identical in terms of all shocks and in both of them all females meet the first partner at $\bar{t}$, but in $\mathcal{S}_{m,p}$ they get an unexpected fertility shock at $\bar{t}$ and in $\mathcal{S}_{m,np}$ they do not. After $\bar{t}$ the sequences are identical (including chances of meeting other partners if the first marriage does not happen/breaks up).  For an auxiliary considerations I also generate sequence of shocks $\mathcal{S}_{nm}$, where at $\bar{t}$ (and before) the agents did not meet anyone.

First, one can look at average impact of the shock, presented at Figure \ref{individual-av}. I take $\bar{t} = 25$. On the top panel, I present probabilities of being married for those who agreed to marry to the partner they met at 25 (so at 25 this is one). As we can see, In the middle, there are unconditional shares of married for two groups. In the bottom I present average utility surplus (love shock) of married couples.

There are two important observations. First, conditional on having satisfactory partner women who did receive fertility shocks have around higher chances of not being married (to anyone) in 10 years. Second, threshold for satisfactory partners falls upon receiving fertility shock: more people marry if the shock happens and average surplus of married couples is lower. This indicates the presence of females who married \emph{because} the pregnancy shock happened (meaning that they would reject the partner if they were not pregnant). Then we can notice that after a number of years surviving couples become very identical in the amount of ``love shock'', meaning that couples with low $\psi$ break up. Third, in the model the largest impact of fertility shock is hurting future marriage prospective for those who stayed a single mother. This is curious although pretty speculative, as the lifecycle of single mothers is modeled in a pretty stylized way.

In the second part, I decompose the observable impact on three groups: those who would marry the partners they met with and without fertility shocks (always-marry), those who marry only if fertility shocks happen (compliers) and those who never marry their current matches (never-marry). In principle, there could exist a group of those who marry only without the shock (defiers), but it is empty in my simulations. Figure \ref{ind-decomp} shows this decomposition. Two dashed lines correspond to the top panel of Figure \ref{individual-av}. The average impact of shock (dashed line with crosses) is a weighted average of two groups (presented by two solid lines). We can see that mainly the divorce probability is driven by small (8.7\%) group of compliers who have dramatically lower probabilities to stay married. Effect on fertility shocks on those who would've married anyway still persists, but is around 1\%. %(this is, however, substantial if we compare it to average probability of not-staying-married without shock that is slightly above 10\%.

%In terms of characteristics, the group of compliers have about 14\% higher median savings and 1.7\% higher median labor income relative to the general population. What makes them compliers is mainly the value of love shock $\psi$: for this group, it is lower than the shock for general population of new couples by about 2/3 of initial standard deviation (i.e. $\sigma(\psi_0)$). This sort of highlights that  the compliers are generally more picky group (that is roughly consistent with evidence on education/income partition) who got a pregnancy shock together with unlucky draw of $\psi$. This is however partially hardcoded in the model, as I target different divorce probabilities for higher-income women.

\begin{figure}[h!]
\centering
\includegraphics[width=0.85\linewidth]{individual-av.eps}
\vspace{-3em}
\caption{Average impact of fertility shock at $t=25$\label{individual-av}}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.85\linewidth]{ind-decomp.eps}
\vspace{-0.5em}
\caption{Decomposition of the average impact\label{ind-decomp}}
\end{figure}

\subsubsection{Aggregate Impact}
I consider two counterfactual scenarios: when unexpected fertility never happens at the moment people meet and when it never happens at all (but people still can make kids by choice) and when there are no fertility completely. I compare key targets in Table \ref{cf-shocks}.The key insight is that fertility shocks do look as a sizable contributor to observable divorce rates, especially at early ages and for lower income couples. This kind of aligns with intuitive understanding of how unplanned pregnancies affect households, and the whole magnitude is roughly in line with vairation of share of K$\to$M couples.

%First, absence of fertility shocks reduces early fertility (in the model every couple before 37 can have a baby by choice). Then, fertility shocks at meeting are sizable driver of observable divorce rates, especially for higher-income couples and early divorces. Finally, fertility shocks in couples (i.e. mistimed births) are also large contributor to the divorce rate,disabling them eliminates early divorces almost completely (though is not so important for later divorces relative to shocks at births). 
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|}\hline
Target & Baseline & No shocks at meeting & No shocks at all \\\hline
\multicolumn{4}{|c|}{Marriage Dimension} \\\hline
Never married by 25  & $60.6\%$ & $60.7\%$ & $60.7\%$  \\
Never married by 35 & $24.6\%$ & $24.3\%$ & $24.3\%$  \\
Never married by 40 & $19.0\%$ & $18.8\%$ & $18.8\%$  \\\hline
\multicolumn{4}{|c|}{Fertility Dimension} \\\hline
No kids at 25 & $65.0\%$ & $68.4\%$ & $68.9\%$ \\
No kids at 30 (low $30\%$ income) & $52.1\%$  &   $57.5\%$ & $59.3\%$ \\
No kids at 30 (top $30\%$ income) & $56.8\%$  &   $57.7\%$ & $58.0\%$ \\
No kids at 35  & $27.2\%$ &   $28.8\%$ & $29.3\%$ \\
No kids at 40 & $23.8\%$  &  $25.0\%$  & $25.4\%$ \\\hline
\multicolumn{4}{|c|}{Divorce Dimension} \\\hline
Divorced by 25 & $4.5\%$   &  $3.2\%$ & $2.3\%$ \\
Divorced by 30 (low $30\%$ income)  & $17.2\%$  &     $14.6\%$ & $11.9\%$ \\
Divorced by 30 (top $30\%$ income)  & $9.0\%$ &     $8.3\%$  & $7.4\%$ \\
Divorced by 35  & $10.8\%$  &   $9.9\%$ & $9.2\%$  \\
Divorced by 40  & $16.3\%$ &  $15.4\%$ & $14.8\%$ \\\hline
\end{tabular}
\caption{Contribution of fertility shocks \label{cf-shocks}}
\end{center}
\end{table}

\section{Conclusions [TBD, these are just some thoughts for the future section]}
There are several qualitative conclusions to be drawn. First, the story of K$\to$M couples experiencing unexpected fertility shocks (rather than choosing to have kids while being unmarried) aligns with the data and economic theory reasonably well. The data presents them as having lower quality marriages, and the theory incorporating bargaining provides a good rationale for why this happens. In this sense, dynamic bargaining theories get another validation by rationalizing this fact.

%Second, aggregate impact of pregnancy shocks is relatively larger for younger and lower income couples, while impact of selection into marriage is more pronounced for richer households. 
Second, the data highlight the importance of time-money substitutability: for instance, in counerfactual simulations when the kids are born being adult, K$\to$M couples still divorce more because unexpected pregnancies create distress, but the gradient in terms of income cannot be seen. In fact, this highlights the idea that less productive couples can mitigate the distress caused by shock by substituting money with time, where higher-income couples do not have this option. This emphasizes importance of time use in modeling childcare.
\clearpage

\section*{Literature}
\bibliography{bib}

\newpage
\appendix
\section{Additional Empirical Results\label{extra-comparisons}}
This part shows few additional comparisons supplemental to the numbers I present in the main part.

\subsection{Differences in Observables\label{comp-diff-appendix}}
This section captures demographic and economic differences of K$\to$M and M$\to$K groups. Many factors could compute to the observable differences, here I quantify them and this justifies use of controls in the next part. I present these differences for ACS, SIPP data are not very different except smaller sample size. For ACS, since M$\to$K and K$\to$M are defined only for subsample of people who were married once and have children present, I also compare this subsample with more general female population. See Table \ref{diff-contr} for the results.

\begin{table}[h!]
\begin{center}\small
\begin{tabular}{|l|c|c|c|c|}\cline{2-5}
\multicolumn{1}{c|}{} & \multirow{2}{*}{All population} & \multicolumn{3}{c|}{Married once + have kids}\\\cline{3-5}
\multicolumn{1}{c|}{}  & & Everyone & K$\to$M & M$\to$K \\\cline{2-5}\hline
\textit{Age} & 30.2    &  33.1     &       31.7     &       33.6 \\
\textit{Age at first marriage} & 24.2         &   23.3        &    23.6        &    23.2 \\
\textit{Age at first birth}  & 24.0       &     25.4     &       21.7       &     26.5 \\
\textit{Husband's age} & 34.9        &    35.8      &      34.4   &         36.2  \\
\textit{Age difference (W$-$H)} &  2.9       &      2.8      &       3.0      &       2.7  \\\hline\hline
\textit{Eldest child's age} & 8.6         &    7.8    &        10.0     &        7.1 \\
\textit{Number of children} & 1.1       &      2.0      &       2.3     &       2.0 \\\hline\hline
\textit{\% college or + (own)} & 32.1\%   &         39.5\%     &       16.9\%     &       46.3\% \\
\textit{\% college or + (husband)} & 33.3\%     &       36.4\%     &       14.2\%       &     42.7\% \\\hline\hline
\textit{\% In Labor Force} & 75.6\%    &        68.6\%    &        70.0\%        &    68.2\% \\
\textit{Own income (thousands)} & 32.2 &         38.4 &          29.1 &         41.1 \\
\textit{Hours worked} & 36.6       &     36.5    &        36.8        &    36.4\\\hline\hline
\textit{\% white (incl hispanic)} & 69.2\%      &      76.5\%      &      72.2\%      &      77.8\% \\
\textit{\% black} & 14.8\% &        6.7\% &            12.2\%       &       5.0\% \\\hline
\multicolumn{5}{|p{0.7\linewidth}|}{\footnotesize ACS, women, 20--40. Means. Everything is subsamples when applicable, income and hours are for subsample of women who are employed.}\\\hline
\end{tabular}
\caption{Differences in composition, ACS\label{diff-comp}}
\end{center}
\end{table}

Shortly, differences in composition are pretty drastic for income and education dimension (as suggested by different shares of K$\to$M in subgroups described above). Most of these things are intuitive, and this suggests importance of flexibly controlling for composition. Similar patterns are observed if I compare composition within each subgroup that I discuss above, so partitions should matter.

\subsection{Duration Considerations}
One obvious concern about comparing two groups is the temporal nature of the data: the variable I use captures fact that divorce happened in the past, and the exact timing may be different between the two groups. Due to cumulative nature of the divorce indicator, females who had marriage long time ago are more likely to be divorced even if their divorce chances in any given period are the same. Therefore, I it is useful to look at divorced females conditional on number of years passed from marriage, see the left panel of Figure \ref{pic-aftmar} for this.
\begin{figure}[h!]
\includegraphics[width=0.45\linewidth]{div_by_aftmar.eps}
\includegraphics[width=0.45\linewidth]{aftmar_by_mk.eps}
\caption{Divorced by years after marriage and distribution of years after marriage\label{pic-aftmar}}
\end{figure}

First, the pattern about different divorce probabilities hold almost uniformly by the ``age of marriage''. Group that has kids first both have higher share of divorced and faster growth of it as marriage progresses. Combining these two graphs allows to see that M$\to$K group are on average married for longer time at the moment of survey, but overall composition is not dramatically different. That would mean that comparing raw shares of divorced actually \emph{underestimates} the difference relative to controlling for number of years after marriage. This is confirmed by regression results that I skip, as in the following section I present the version where I control for years after marriage and few other factors.

\subsection{Effects by Child's Age}
First, I show how share of K$\to$M people varies with age of the eldest child. I focus on subsample children are 18 or younger, as very few people have children above 18 living in the household. This exercise highlights a nuance in the construction of the groups: almost no one in K$\to$M group has children of age less than a year, because that would mean the same year of marriage, first birth and ACS survey. Additionally, since the population is trimmed at the age of 40, having kids around age of 18 means fairly low age of the first birth, that is typically associated with being in M$\to$K group. Left panel of Figure \ref{pics-raw} clearly depicts both issues, and this suggests the importance of using controls. In the specifications above I control both for number of years after marriage and the age of the eldest child, so this should help mitigating these issues.

Right panel of Figure \ref{pics-raw} shows evolution of divorce probabilities for the two groups. The increasing pattern is a result of cumulative nature of processes: people with elder children have more periods to get divorced. But the difference between two probabilities is illustrative: between ages 5 and 10 size of K$\to$M group changes a little, but difference becomes very pronounced.

\begin{figure}[h!]
\includegraphics[width=0.45\linewidth]{km_by_age.eps}
\includegraphics[width=0.45\linewidth]{div_by_age_raw.eps}
\caption{Share of K$\to$M couples and share of divorced, no controls\label{pics-raw}}
\end{figure}

I focus on studying difference more carefully then: first, I simply run a regression
\[D_i = \sum_{a=0}^{18} \beta_a \cdot \I(\text{Age of eldest child}_i = a) +  \sum_{a=0}^{18} \Delta_a \cdot \I(\text{Age of eldest child} = a)\cdot \I(\text{K$\to$M})_i + \varepsilon_i,\]
and plot $\Delta_a$.

Second, for a subsample of working women I run a regression with controls that were used for Table \ref{diff-contr} above (including polynomials in income),
\[D_i = \sum_{a=0}^{18} \Delta_a \cdot \I(\text{Age of eldest child} = a)\cdot \I(\text{K$\to$M})_i + X_i\gamma + \varepsilon_i,\]
note that these controls include dummies for the age of the eldest child.

See Figure \ref{pics-diff} for the results. The left panel essentially replicates the differences on the previous graph. The right panel is of the main focus: if the set of controls captures the issues of construction of the groups, then the differences associated with age follow the conjecture above: they are small for younger children and start to hike at ages 5--10 (I believe that first few points are mostly caused by composition). In general, these graphs are kind of consistent with the idea that divorces are concentrated around the moments where kids get older and childcare costs structure changes as they arguably require less time, though they do not provide a formal test.
\begin{figure}[h!]
\includegraphics[width=0.45\linewidth]{diff_raw_by_age.eps}
\includegraphics[width=0.45\linewidth]{diff_cont_by_age.eps}
\caption{Differences in divorce probabilities by child's age\label{pics-diff}}
\end{figure}

\subsection{Sample of Recently Divorced}
In this exercise I utilize an alternative variable available in ACS --- indicator whether a particular person got a divorce in the year of survey. It can be interpreted as \textit{marginal} divorce probability as opposed to the previous value being \emph{cumulative}. Therefore, for instance, it does not have an upward trend in number of years married, see Figure \ref{pic-now-aftmar}. 
\begin{figure}[h]
\centering
\includegraphics[width=0.45\linewidth]{divnow_by_aftmar.eps}
\caption{Probabilities of getting divorced in the year of survey\label{pic-now-aftmar}}
\end{figure}

Table \ref{diff-contr-marginal} replicates the results for marginal divorce probabilities. The key conclusions and proportions remain very similar, that implicitly suggests validity of the approach used above.

\begin{table}
\begin{center}
\begin{tabular}{|l|c||c|c|c|c|c|c|}\cline{2-8}
\multicolumn{1}{c|}{} & \multicolumn{7}{c|}{ACS: sample of married once}\\\cline{2-8}
\multicolumn{1}{c|}{} & \small Share of &\multicolumn{2}{c|}{\small Raw difference} & \multicolumn{2}{c|}{\small with controls} & \multicolumn{2}{c|}{\small controls + income} \\ \cline{3-8}
\multicolumn{1}{c|}{} & \small divorced & $\Delta$ & (s.e.) & $\Delta$ & (s.e.) & $\Delta$ & (s.e.) \\  \hline
\textit{All females, 20--50} & 1.94\% &  1.37\% &  (0.05\%) &  0.82\% &  (0.08\%) &  0.79\%  &  (0.11\%) \\
\textit{All females, 35--40} &   1.59\% &  1.04\% &  (0.08\%) &   0.77\% &  (0.11\%)  &  0.74\%   &  (0.15\%) \\\hline\hline
\multicolumn{8}{|p{0.9\linewidth}|}{\footnotesize \textbf{Education partition.}}\\\hline
\textit{High school only, 20--50} & 2.31\% &  0.86\% &  (0.08\%) &   0.82\% &  (0.12\%)  &  0.82\%  &   (0.20\%) \\
\textit{College graduates, 20--50} & 1.19\% &  1.24\% & (0.10\%) &   0.59\% &   (0.13\%)  &  0.50\% &    (0.16\%) \\\hline\hline
\multicolumn{8}{|p{0.9\linewidth}|}{\footnotesize \textbf{Income partition:} working women only.}\\\hline
\textit{Low 30\% income, 20--40} &  3.02\% &  1.31\% &  (0.14\%) &   0.90\% &  (0.21\%) &   0.91\%  &   (0.21\%) \\
\textit{Mid 40\% income, 20--40} &  2.54\% &  1.34\% &  (0.12\%) &   0.54\%  & (0.18\%) &   0.54\% &    (0.18\%) \\
\textit{Top 30\% income, 20--40} & 1.69\% &  1.69\% &   (0.15\%) &   0.90\%  & (0.20\%) &   0.90\% &    (0.20\%) \\
\hline
\end{tabular}
\caption{Differences in marginal probablities, controlled for composition, ACS\label{diff-contr-marginal}}
\end{center}
\end{table}

\subsection{Step or Own\label{step-part}}
People with step-children ex-ante have different marriage characteristics. However, as I restrict the sample to $\Delta T \geq 0$, the share of stepchildren in the observable couples is about $14\%$. However, this number is not perfectly observed, as ACS reports only relationships of children to the household's head, so whether the children are step can only be seen for existing couples where the head is male. Table \ref{step-1} presents total counts of households by values of $\Delta T$ who report that their eldest child is step, as you can see, it sharply declines with $\Delta T$ approaching zero and total share is slightly above one-eights. It also shows average years of being married at the moment of survey, and if couples with step children were breaking up more quickly this duration would have been lower for step group, which is not true in the data.
\begin{table}
\centering
\begin{tabular}{|r|r|r||r|r|}\hline
& \multicolumn{2}{c||}{Total count} & \multicolumn{2}{c|}{Years married} \\\hline
 $\Delta T$ &  Own & Step & Own & Step\\\hline
        $-5$ &     6,676  &     2,776 & 6.07  & 5.94\\
        $-4$ &     7,952  &    2,907 & 6.48 &  6.49  \\
        $-3$ &     9,513  &    2,638 & 6.85 & 6.87\\
        $-2$ &    11,905 &     2,227 & 7.13   & 7.24\\
        $-1$ &     15,344 &    1,390 &  7.52  & 7.81 \\
         $0$  &  28,657  &      715  & 8.27  & 8.81 \\\hline
     All $\leq 0$ &    80,047 &    12,653 \\
     (shares)        & ($86.4\%$)  & ($13.6\%$)\\\cline{1-3}
\end{tabular}
\caption{Counts and marriage durations (at the moment of survey) of K$\to$M households with step and own children, ACS.\label{step-1}}
\end{table}

\input{theory-appendix} % notes on model
\input{comp-appendix} % notes on computations
\input{tech-appendix} % notes on solution process

\end{document}